Automated Formal Synthesis\\ of Safe Digital Controllers for Continuous Plants

### 2018-05-16 automatica/paper.tex
intro

\begin{keyword}                           % Five to ten keywords,  
Digital control synthesis; Safety requirements; Analog/Digital conversion; Quantization errors; Sampling.               % chosen from the IFAC 
\end{keyword}                             % keyword list or with the 
                                          % help of the Automatica 
                                          % keyword wizard

@@

Our work addresses challenging aspects of the control synthesis problem:  
we perform digital control synthesis over a ``hybrid model'',
namely a model encompassing a plant exhibiting continuous behavior, 
whereas the controller operates in discrete time and over a quantized domain.  
In particular, our model evaluates the effects of the quantizers (A/D and D/A
converters) and of time discretization, as well as representation
errors introduced by working  %% artifacts such as an
%% observer~\cite{astrom1997computer}
in a finite-precision domain. 
We reason about high level properties such as stability, robustness, and safety, which define the
behaviour of the whole model.  
Notably, safety requirements are frequently overlooked in conventional feedback control synthesis, but
play a key role in systems engineering, particularly in safety-critical applications.

@@

Instead of unfolding the dynamics, \emph{the second approach}
employs \emph{abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the model simultaneously. 
Additionally, this approach uses \emph{abstraction refinement} \cite{DBLP:conf/cav/ClarkeGJLV00},
enabling one to start with a very simple (abstract) description of the model 
regardless of the complexity of its dynamics, and to refine to more complex models only 
when a solution cannot be found.

@@

%%   \note{[this statement is way too broad] Existing methods for controller synthesis rely on transfer 
%%   function representations, which are inadequate to prove safety requirements. } 
%% %
\item We provide two novel algorithms: the first, a multi-staged approach, relies on 
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction/refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of 
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant. 
%
\item We develop a model for different sources of quantization error and account for 
  their effect on safety.  We give bounds that ensure the
  safety of our controllers in a hybrid (i.e., continuous/digital) domain.

\item While~\cite{DBLP:conf/cav/AbateBCCDKKP17} is restricted to models in discrete 
time, in the current paper we present an approach for
  soundly evaluating safety in combined continuous-discrete time and space.

\item A further limitation of the work in~\cite{DBLP:conf/cav/AbateBCCDKKP17} 
  is its restriction to fixed-point arithmetic, meaning that fixed-point
  numbers are employed for the representation of both the plant and the
  controller, as well as for the operations performed by each of them. 
  Conversely, in the current paper we also make use of floating-point
  arithmetic, and compare the two approaches.  Additionally, we enhance the
  implementation with an SMT back-end, allowing us to also use reals in order
  to represent the states of the plant.

@@

The work in~\cite{hscc-paper} synthesizes stabilizing controllers for
continuous plants given as transfer functions by exploiting bit-accurate
verification of software-implemented digital controllers~\cite{Bessa16}. 
While this work also uses CEGIS, the approach is restricted to digital
controllers for stable closed-loop systems given as transfer function
models: this results in a static check on their coefficients.  By contrast,
in the current paper we consider a state-space representation of the
physical system, which requires ensuring the specification over actual
traces of the model, alongside the numerical soundness required by the
effects of discretisation and finite-precision errors.
\note{Next paragraph is naive for a control audience - amend} 
A~state-space model has known advantages over the transfer function
representation~\cite{Franklin15}, as 
% it naturally generalizes to multivariate systems (i.e., with multiple inputs and outputs); and 
it allows synthesis of controllers with guarantees on the internal dynamics, e.g., 
%to synthesize controllers that make the closed-loop model 
\emph{safety}.  Our work indeed focuses on the \emph{safety} of internal states, which is usually
overlooked in the standard control literature with its default focus on stability properties.  
Moreover, our work integrates an abstraction/refinement (CEGAR) step inside the main CEGIS loop.

The approach underpinning tools such as Pessoa~\cite{mazo2010pessoa} and related extensions \cite{Anta2010,liu16} and applications \cite{zamani2014}, 
synthesizes correct-by-design embedded control software for LTI models.    
It is based on the abstraction of the model to a finite-state machine, 
and on the formal control synthesis over safety- and reachability-based temporal properties thereon.  
%Based on this safety specification, \mbox{Pessoa} can synthesize embedded controller software for a range of
%properties.  
The controller software can then be refined (implemented) over the concrete LTI model. 
%can be more complicated than
%the state-feedback control we synthesize, and the properties available cover more detail.  
However, relying on state-space discretization \mbox{Pessoa} is likely to incur in scalability limitations.   
%Along this research line, \cite{Anta2010,liu16} studies the synthesis of digital controllers for
%continuous dynamics, and \cite{zamani2014} extends the approach to the Network Control Systems setup.

>>>>

\begin{keyword}                           % Five to ten keywords,  
Digital control synthesis; Safety requirements; Analog/Digital conversion; Quantization errors; Time sampling.               % chosen from the IFAC 
\end{keyword}                             % keyword list or with the 
                                          % help of the Automatica 
                                          % keyword wizard

@@

Our work addresses challenging aspects of the control synthesis problem:  
we perform digital control synthesis over a ``hybrid model'',
namely a model encompassing a plant exhibiting continuous behavior, 
whereas the controller operates in discrete time and over a quantized domain.  
In particular, our model evaluates the effects of the quantizers (A/D and D/A
converters) and of time discretization, as well as representation
errors introduced by working  %% artifacts such as an
%% observer~\cite{astrom1997computer}
in a finite-precision domain. 
We reason about high level properties such as stability, robustness, and safety, which define the
behaviour of the whole model.  
Notably, safety requirements are frequently overlooked in conventional feedback control synthesis, but
play a key role in systems engineering, particularly for safety-critical applications of clear concern in modern autonomy contexts.

@@

Instead of unfolding the dynamics, \emph{the second approach}
employs \emph{abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the model simultaneously. 
Additionally, this approach uses the \emph{abstraction-refinement} template \cite{DBLP:conf/cav/ClarkeGJLV00},
enabling one to start with a very simple (abstract) description of the model 
regardless of the complexity of its dynamics, and to refine to more complex models only 
when a solution cannot be found.

@@

%%   \note{[this statement is way too broad] Existing methods for controller synthesis rely on transfer 
%%   function representations, which are inadequate to prove safety requirements. } 
%% %
\item We provide two novel algorithms: the first, a multi-staged approach, relies on 
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction/refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of 
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant. 
%
\item We develop a model for different sources of quantization error and account for 
  their effect on safety.  We provide bounds that ensure the
  safety of our controllers in a hybrid (i.e., continuous/digital) domain.

\item While~\cite{DBLP:conf/cav/AbateBCCDKKP17} is restricted to models in discrete 
time, \note{in the current paper we present an approach for
  soundly evaluating safety in combined continuous-discrete time and space. [this is via the AA-based back-end. However this part is not in the experiments, afaik.]}

\item A further limitation of the work in~\cite{DBLP:conf/cav/AbateBCCDKKP17} 
  is its restriction to fixed-point arithmetic, meaning that fixed-point
  numbers are employed for the representation of both the plant and the
  controller, as well as for the operations performed by each of them. 
  Conversely, in the current paper we also make use of floating-point
  arithmetic, and compare the two approaches.  \note{Additionally, we enhance the
  implementation with an SMT back-end, allowing us to also use reals in order
  to represent the states of the plant.}

\item \note{Shall we mention the broader set of benchmarks?}

@@

The work in~\cite{hscc-paper} synthesizes stabilizing controllers for
continuous plants given as transfer functions by exploiting bit-accurate
verification of software-implemented digital controllers~\cite{Bessa16}. 
While this work also uses CEGIS, the approach is restricted to digital
controllers for stable closed-loop systems given as transfer function
models: this results in a static check on their coefficients.  By contrast,
in the current paper we consider a state-space representation of the
physical system, which requires ensuring the specification over actual
traces of the model, alongside the numerical soundness required by the
effects of discretisation and finite-precision errors.
\note{Next paragraph is naive for a control audience - amend} 
A~state-space model has known advantages over the transfer function
representation~\cite{Franklin15}, as 
% it naturally generalizes to multivariate systems (i.e., with multiple inputs and outputs); and 
it allows synthesis of controllers with guarantees on the internal dynamics, e.g., 
%to synthesize controllers that make the closed-loop model 
\emph{safety}.  Our work indeed focuses on the safety of internal states, 
which is by and large overlooked in the standard (digital) control literature, 
by default focussed on stability properties.  
Moreover, our work integrates an abstraction-refinement (CEGAR) step inside the main CEGIS loop.

The technique underpinning tools such as Pessoa~\cite{mazo2010pessoa} and related extensions \cite{Anta2010,liu16} and applications \cite{zamani2014}, 
synthesizes correct-by-design embedded control software for LTI models.    
It is based on the abstraction of the model to a finite-state machine, 
and on the formal control synthesis over safety- and reachability-based temporal properties thereon.  
%Based on this safety specification, \mbox{Pessoa} can synthesize embedded controller software for a range of
%properties.  
The controller software can then be refined (implemented) over the concrete LTI model. 
%can be more complicated than
%the state-feedback control we synthesize, and the properties available cover more detail.  
However, relying on state-space discretization \mbox{Pessoa} is likely to incur in scalability limitations.   
%Along this research line, \cite{Anta2010,liu16} studies the synthesis of digital controllers for
%continuous dynamics, and \cite{zamani2014} extends the approach to the Network Control Systems setup.
